# Mixed effects analyses for colexification data

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=4, warning=FALSE, message=FALSE)
```

### Libraries

```{r libraries}
library("lme4")
library("R.matlab")
```

### Data

```{r datafiles}
datafiles <- c("rdictisdata.mat", "ridsplusdata.mat")
datanames <- c("dictionary", "idsplus")
# index of ice/snow in each data set (ridsplusdata.mat includes colexification data for many pairs)
icesnowinds <- c(1, 15043)
# index of mean monthly temperature in data set
climind <- 6 
# mean temperature for English (stan1293)
engmean <- 8.68333
# mean temperature for Nahuatl (high1278)
nahmean <- 21.875 
```

### Analysis loop

```{r analysis}

logit2prob=function(L){
  exp(L)/(1+exp(L))
}

for (i in 1:2)  {
  # create data frame
  alldata <- readMat(datafiles[i])
  dataname <- datanames[i]
  icesnowind <- icesnowinds[i]
  sameword <- alldata$reditd[,icesnowind]
  temp <-alldata$rorigclimdata[,climind]/10
  langnames <- sapply(alldata$rlabels, function(n) n[[1]])
  langfamily <- alldata$rfamcodes
  area <- alldata$rautotypcodes
  d <- data.frame(sameword, temp, langfamily, area)

  # drop cases with missing data
  completeind <- complete.cases(d)
  andata <- d[completeind,]

  nlump<- sum(andata$sameword==1)
  nsplit<- sum(andata$sameword==0)

  cat(c("# lumpers: ", as.character(nlump), "\n"))
  cat(c("# splitters: ", as.character(nsplit), "\n"))


  # box and whisker plot
  boxplot(temp ~ sameword, data = andata, names = c("split", "merge"), horizontal = TRUE)
  title(dataname, xlab = "mean temperature")

  # run mixed effects analysis
  m1 <- glmer(sameword ~ temp + (1|langfamily), data = andata, family = binomial) 
  m2 <- glmer(sameword ~ (1|langfamily), data = andata, family = binomial)
  a <- anova(m2, m1)


  # Effect size
  # probability of same word at temperature associated with English
  prob.at.t0 <- logit2prob(fixef(m1)[1]+ (fixef(m1)[2]*engmean))
  # probability of smae word at temperature assciated with Nahuatl
  prob.at.t1 <- logit2prob(fixef(m1)[1] + (fixef(m1)[2]*nahmean))
  cat("Prob of Same word at t0 (English) =",  sprintf("%.3f\n", prob.at.t0))
  cat("Prob of Same word at t1 (Nahuatl) =",  sprintf("%.3f\n", prob.at.t1))

  #  summarize results
  cat(c("Summary of", dataname, "analysis:\n"))
  print(summary(m1))
  cat("\n\n\n\n")
  cat(c("Likelihood ratio test for", dataname, "data :\n"))
  print(a)
  cat("\n\n")

  # should we add area random effect?
  print("Add Random effect for area?")
  m2_area <- glmer(sameword ~ (1|langfamily) + (1|area), data = andata, family = binomial)
  print(anova(m2,m2_area))

  # with area random effect
  print("With Area Random effect")
  m1.area <- glmer(sameword ~ temp + (1|langfamily) + (1 |area), data = andata, family = binomial) 
  m2.area <- glmer(sameword ~ (1|langfamily) + (1 |area), data = andata, family = binomial)
  print(summary(m1.area))
  print(anova(m2.area, m1.area))

  print("Slope by area")
  m1.areaslope <- glmer(sameword ~ temp + (1|langfamily) + (1 |area) + (temp|area), data = andata, family = binomial, control=glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)) ) 
  m2.areaslope<- glmer(sameword ~ (1|langfamily) + (1 |area) + (temp|area), data = andata, family = binomial, control=glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000)) )
  print(summary(m1.areaslope))
  print(anova(m2.areaslope, m1.areaslope))

  # probability of same word at temperature associated with English
  prob.at.t0 <- logit2prob(fixef(m1.areaslope)[1]+ (fixef(m1.areaslope)[2]*engmean))
  # probability of smae word at temperature assciated with Nahuatl
  prob.at.t1 <- logit2prob(fixef(m1.areaslope)[1]+ (fixef(m1.areaslope)[2]*nahmean))
  cat("Prob of Same word at t0 (English) =",  sprintf("%.3f\n", prob.at.t0))
  cat("Prob of Same word at t1 (Nahuatl) =",  sprintf("%.3f\n", prob.at.t1))
}
