# Mixed-effects analysis of missing forms for ice and/or snow

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=4, warning=FALSE, message=FALSE)
```

### Libraries

```{r libraries}
library("lme4")
library("R.matlab")
```

### Data

```{r datafiles}
datafiles <- c("ridsplusdata.mat")
datanames <- c("idsplus")
# index of ice/snow in each data set (ridsplusdata.mat includes colexification data for many pairs)
icesnowinds <- c(15043)
# index of mean monthly temperature in data set
climind <- 6 
```

### Analysis loop

```{r analysis}
for (i in 1:1)  {
  # create data frame
  alldata <- readMat(datafiles[i])
  dataname <- datanames[i]
  icesnowind <- icesnowinds[i]
  sameword <- alldata$reditd[,icesnowind]
  missform<- sameword
  missform[is.nan(sameword)] <- 1
  missform[!is.nan(sameword)] <- 0
  temp <-alldata$rorigclimdata[,climind]/10
  langnames <- sapply(alldata$rlabels, function(n) n[[1]])
  langfamily <- alldata$rfamcodes
  langmat <- alldata$rlangmat
  d <- data.frame(missform, temp, langfamily)
  # include only languages with forms for 1000 or more meanings
  d <- d[rowSums(langmat)>=1000,]

  # drop cases with missing data
  completeind <- complete.cases(d)
  andata <- d[completeind,]
  nmiss<- sum(andata$missform==1)
  nnomiss<- sum(andata$missform==0)

  cat(c("# missing form or snow and/or ice: ", as.character(nmiss), "\n"))
  cat(c("# without missing forms: ", as.character(nnomiss), "\n"))

  # box and whisker plot
  boxplot(temp ~ missform, data = andata, names = c("both present", "missing"), horizontal = TRUE)
  title(dataname, xlab = "mean temperature")

  # run mixed effects analysis
  m1 <- glmer(missform ~ temp + (1|langfamily), data = andata, family = binomial) 
  m2 <- glmer(missform ~ (1|langfamily), data = andata, family = binomial)
  a <- anova(m2, m1)

  #  summarize results
  cat(c("Summary of", dataname, "analysis:\n"))
  print(summary(m1))
  cat("\n\n\n\n")
  cat(c("Likelihood ratio test for", dataname, "data :\n"))
  print(a)
  cat("\n\n")
}
