# Logistic regression analysis for USA English

```{r knitr_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=6, fig.height=6, warning=FALSE, message=FALSE)
```

### Libraries

```{r libraries}
library("lme4")
library("R.matlab")
```

### Data

```{r datafiles}
datafiles <- c("../freqdata/rtwitterusa.mat")
datanames <- c("Twitter (USA)")
```

### Analysis loop

```{r analysis}
for (i in 1:1)  {
  # create data frame
  alldata <- readMat(datafiles[i])
  dataname <- datanames[i]
  temp <-alldata$rorigclim/10
  # number of ice/snow tokens per language
  iscount<- alldata$rrcount[,1]
  # number of tokens per language
  allcount<- alldata$rrcount[,2]
  langnames <- sapply(alldata$rlabels, function(n) n[[1]])
  d <- data.frame(iscount, allcount, temp)

  # drop cases with missing data
  completeind <- complete.cases(d)
  andata <- d[completeind,]

  # make scatter plot
  plot(temp, log(iscount/(allcount-iscount)), xlab = "temperature", ylab = 'log odds')
  title(dataname)

  # run logistic regression
  m1 <- glm(cbind(iscount, allcount - iscount) ~ temp, data = andata, family = binomial)
  m2 <- glm(cbind(iscount, allcount - iscount) ~ 1, data = andata, family = binomial)
  a <- anova(m2, m1, test="LRT")

  #  summarize results
  cat(c("Summary of", dataname, "analysis:\n"))
  print(summary(m1))
  cat("\n\n\n\n")
  cat(c("Likelihood ratio test for", dataname, "data :\n"))
  print(a)
  cat("\n\n")
}

